---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='figure/',
                      warning=FALSE, message=FALSE)
```

## Step 1. Loading and preprocessing the data

The variables included in research dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)
* **date**: The date on which the measurement was taken in YYYY-MM-DD format
* **interval**: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file (activity.csv) and there are a total of 17,568 observations in this dataset. CSV file is packed as zip. The following code unpacks the file in data folder and loads in "data" variable.


```{r}
  if (!file.exists("./data")) {
    dir.create("./data");
  }
  if (!file.exists("./data/activity.csv")) {
    unzip("activity.zip", exdir = "data");
  }  
  data <- read.csv("./data/activity.csv")

```

Number of rows loaded:
```{r echo=FALSE}

nrow(data)

```


## Step 2. What is mean total number of steps taken per day?

The following code calculates total number if steps per each day

```{r}
  stepsPerDay <- aggregate(. ~ date, data = data, sum)
  stepsPerDay[, c("date", "steps")]

```

The disribution of number of steps represented by the following histogram:

```{r steps_per_day_hist, warning=FALSE}
  library(ggplot2)
  ggplot(stepsPerDay, aes(x=steps)) +
    geom_histogram(colour="black", fill="orange")
```

Mean and Median of steps per day are the following
```{r}
  mean(stepsPerDay$steps)
  median(stepsPerDay$steps)
```


## Step 3. What is the average daily activity pattern?

Let's calculate averages of steps for each time interval over all dates:

```{r}
  intraday <- aggregate(. ~ interval, data= data, mean)[,c("interval", "steps")]
```

The following graph shows distribution of number of steps over the day
```{r steps_per_5minutes_dist, warning=FALSE}
  ggplot(intraday, aes(x=interval, y=steps)) + geom_line(colour="orange")
```

Maximum number of steps is reached in the following 5-minute interval:
```{r}
  intraday[max(intraday$steps) == intraday$steps, "interval"]

```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
