---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='figure/',
                      warning=FALSE, message=FALSE)
```

## Step 1. Loading and preprocessing the data

The variables included in research dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)
* **date**: The date on which the measurement was taken in YYYY-MM-DD format
* **interval**: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file (activity.csv) and there are a total of 17,568 observations in this dataset. CSV file is packed as zip. The following code unpacks the file in data folder and loads in "data" variable.


```{r}
  if (!file.exists("./data")) {
    dir.create("./data");
  }
  if (!file.exists("./data/activity.csv")) {
    unzip("activity.zip", exdir = "data");
  }  
  data <- read.csv("./data/activity.csv")

```

Number of rows loaded:
```{r echo=FALSE}

nrow(data)

```


## Step 2. What is mean total number of steps taken per day?

The following code calculates total number if steps per each day

```{r}
  stepsPerDay <- aggregate(. ~ date, data = data, sum)
  stepsPerDay[, c("date", "steps")]

```

The disribution of number of steps is represented by the following histogram:

```{r steps_per_day_hist, warning=FALSE}
  library(ggplot2)
  ggplot(stepsPerDay, aes(x=steps)) +
    geom_histogram(colour="black", fill="orange")
```

Mean and Median of steps per day are the following
```{r}
  mean(stepsPerDay$steps)
  median(stepsPerDay$steps)
```


## Step 3. What is the average daily activity pattern?

Let's calculate averages of steps for each time interval over all dates:

```{r}
  intraday <- aggregate(. ~ interval, data= data, mean)[,c("interval", "steps")]
```

The following graph shows distribution of number of steps over the day
```{r steps_per_5minutes_dist, warning=FALSE}
  ggplot(intraday, aes(x=interval, y=steps)) + geom_line(colour="orange")
```

Maximum number of steps is reached in the following 5-minute interval:
```{r}
  intraday[max(intraday$steps) == intraday$steps, "interval"]

```


## Imputing missing values

Number of lines with NA:

```{r}
sum(is.na(data$steps))
```

Let's fill missing values with average of steps for same 5-minute intervals on other day. We will reuse intraday dataset created previously

```{r}
  fulldata <- merge(x=data, y=intraday, by="interval", all.x = TRUE)
  fulldata[is.na(fulldata$steps.x),"steps.x"] <- fulldata[is.na(fulldata$steps.x),"steps.y"]
```

Next we aggregate imputed data by day
```{r}
  stepsPerDayFull <- aggregate(. ~ date, data = fulldata, sum)[, c("date", "steps.x")]

```

The disribution of number of steps with imputed data is represented by the following histogram:

```{r steps_per_day_hist_full, warning=FALSE}
  ggplot(stepsPerDayFull, aes(x=steps.x)) +
    geom_histogram(colour="black", fill="orange")
```

Mean and Median of steps per day is the following are the following
```{r}
  mean(stepsPerDayFull$steps.x)
  median(stepsPerDayFull$steps.x)
```

As result we see that histogram form does not change, but bars height is increased. Als median and mean value are equals now. This is expected cause we use average for 5-minutes interval to impute empty values.

## Are there differences in activity patterns between weekdays and weekends?

Let's fill wday xolumn with weekday/weekend factor

```{r}
fulldata$wday <- factor(ifelse(weekdays(as.Date(fulldata$date)) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
```

Now lets average data by 5-minutes intervals
```{r}
intradayWithDays <- aggregate(steps.x ~ wday + interval, data = fulldata, mean)
```

And display result as graphs
```{r steps_per_day_wday}
ggplot(intradayWithDays, aes(x=interval, y=steps.x)) + 
    geom_line(colour="orange") +
    facet_grid(wday ~ .) +
    ylab("Steps per 5-minutes interval") +
    ggtitle("Step distribution by weekday/weekend")
```
